### 2019-01-21 R/paper 02 Scale-jump analysis of historical retrospective patterns.Rmd
Updates until 20/1/2019

  # Exclude prediction years
  filter(year <= assessmentyear) %>%

  # Define start year
  filter( year >= 1980) %>%

  # Only assessments with stock size in tonnes
  filter(stocksizeunits  == "tonnes") %>%  # Note: this is filtering on absolute vs relative; but not really

@@

  arrange(stock, purpose, desc(assessmentyear)) %>%

  # filter(purpose         == "advice") %>%

  select(assessmentyear, year, stock, stockkeylabelold, purpose, benchmark,
         stocksize, recruitment, fishingpressure) %>% 
  distinct() %>% 
  gather(key=variable, value=value, stocksize:fishingpressure) %>% 
  filter(!is.na(value), value != 0.0)  %>%

@@

  arrange(stock, variable, desc(assessmentyear)) %>% 
  ungroup() %>%

@@

distinct(iAssess, stockkeylabelold, assessmentyear, purpose, benchmark) %>% filter(grepl("ple-n", stockkeylabelold)) %>% View()

@@

# select data
d.selection <-
  d %>%
  filter(purpose == "advice")

# benchmark data
d.benchmarks <-
  d %>% 
  filter(benchmark == TRUE) %>% 
  distinct(stock, stockkeylabelold, assessmentyear, purpose, .keep_all = FALSE)

@@

  left_join(iAdvice, by=c("stockkeylabelold","assessmentyear", "purpose")) %>%
  mutate(assessmentmodel = tolower(assessmentmodel)) %>%

@@

  # determine stock definition change and model definition change
  group_by(stock) %>% 
  arrange(stock, assessmentyear) %>% 
  mutate(assessmentmodelchange = ifelse(assessmentmodel != lag(assessmentmodel), 1, 0),
         assessmentmodelchange = ifelse(is.na(assessmentmodelchange), 0, assessmentmodelchange)) %>%

  group_by(stock) %>% 
  arrange(stock, assessmentyear) %>% 
  mutate(stockdefinitionchange = ifelse(stockkeylabelold != lag(stockkeylabelold), 1, 0),
         stockdefinitionchange = ifelse(is.na(stockdefinitionchange), 0, stockdefinitionchange))

@@

  group_by(stock, variable, calctype) %>% 
  summarize(jonsson_ab   = mean(jonsson_ab),
            jonsson_asd  = sqrt(mean(jonsson_dev2)),
            mohn_rhoabs = mean(mohn_rhoabs),
            mohn_rho     = mean(mohn_rho), 
            firstyear    = min(year),
            lastyear     = max(year))

@@

  # now do summing over years
  mutate(sumlogBdev2 = sum(logBdev2) ) %>%

  # now summarize by year
  summarize(
    meanlogBdev = mean(logBdev),
    sumlogBdev2 = sum(logBdev2),
    n_1         = mean(n_1), 
    ralston_sigma = sqrt(sumlogBdev2/n_1)) %>%

@@

  group_by(stock, calctype) %>%
  distinct(stock, year, n_1, sumlogBdev2, .keep_all = FALSE) %>% 
  summarize(
    sumlogBdev2all = sum(sumlogBdev2),
    n_1            = sum(n_1), 
    firstyear    = min(year),
    lastyear     = max(year)
  ) %>%

@@

# data series with most recent (last) estimation
d.last <-
  d.pairs %>% 
  select(-assessmentyearprior) %>% 
  left_join(d, by=c("stock", "variable", "assessmentyearlast" = "assessmentyear")) %>% 
  mutate(label="last") %>% 
  rename(assessmentyear = assessmentyearlast)

# data series with previous (prior) estimation
d.prior <-
  d.pairs %>% 
  select(-assessmentyearlast) %>% 
  left_join(d, by=c("stock", "variable", "assessmentyearprior" = "assessmentyear")) %>% 
  mutate(label="prior") %>% 
  rename(assessmentyear = assessmentyearprior)

d.jumps <-
  bind_rows(d.last, d.prior) %>% 
  group_by(stock, assessmentpair, variable) %>%

@@

  # filter only years that are in common in last and prior
  group_by(stock, assessmentpair, variable, year) %>% 
  mutate(n = n()) %>% 
  filter(n() == 2) %>%

@@

@@

fig_nums(name    = "scalejumps", level = 1, display = FALSE, 
         caption = "Retrospective scale difference (sigma) for 9 species derived from historical assessment records. Results from pairwise comparisons over periods of 8 years. Large dots  refer to years when the scale difference is larger than 10%. Numbers at the bottom left of each panel refer to the number of assessments with a scale difference larger than 10% relative to the total number of assessments available for that stock. Colours indicate the decades. Black triangles and dashed lines indicate benchmark assessment years. Red triangles and dashed lines indicate years when assessment models were changed compared to the year before")

@@

@@

@@

  ggplot(aes(x=assessmentyear, y=value)) +
  theme_publication() +
  theme(
    panel.spacing    = unit(0.1, "lines"),
    panel.grid.major.x = element_line(colour = "gray80"),
    strip.background = element_blank(),
    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), angle = 180),
    strip.placement = "outside",
    legend.position  = "none"
  ) +

@@

  # geom_point(data=d.benchmarks, 
  #            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +

  # geom_point(data=d.assessmentmodelchange, 
  #            aes(x=assessmentyear-0.1, y=-1.0), colour="red", fill="red", shape=24, size=2) +

  # geom_segment(data=d.benchmarks, 
  #              aes(x=assessmentyear+0.1, xend=assessmentyear+0.1, y=-1.0, yend=1.0), 
  #              colour="gray20", size=0.5, linetype="dashed") +

  # geom_segment(data=d.assessmentmodelchange, 
  #              aes(x=assessmentyear-0.1, xend=assessmentyear-0.1, y=-1.0, yend=1.0), 
  #              colour="red", size=0.5, linetype="dashed") +

@@

  # geom_text(data=filter(d.scale, variable == myvar), 
  #           aes(label=ratio),  
  #           x=-Inf, y=-Inf, hjust=-0.2, vjust=-0.8, inherit.aes=FALSE) +

  coord_cartesian(ylim=c(-1, 1)) +  # set the limits in cartesian space prevents outlier lines to be excluded
  scale_y_continuous(breaks = pretty_breaks(), labels=percent_format()) +
  scale_x_continuous(breaks = pretty_breaks(), limits=c(1982, NA)) +
  labs(x=" ", y="") +
  facet_wrap(~stock)

@@

@@

>>>>

@@

  # Exclude prediction years (but keep the benchmark information that has no year attached to it)
  filter(is.na(year) | (year <= assessmentyear & year >= 1980) ) %>% 
  filter(assessmentyear >= 1982) %>% 
  filter(is.na(year) | !(is.na(stocksize) & is.na(fishingpressure))) %>%

  # Only assessments with stock size in tonnes
  filter(is.na(year) |stocksizeunits  == "tonnes") %>%

  # Sorting
  arrange(stock, purpose, desc(assessmentyear)) %>%

  select(assessmentyear, year, stock, stockkeylabelold, purpose, benchmark,
         stocksize, recruitment, fishingpressure) %>% 
  distinct() %>% 
  gather(key=variable, value=value, stocksize:fishingpressure) %>% 
  filter(is.na(year) | (!is.na(value) & value != 0.0) )  %>%

@@

  arrange(stock, variable, (assessmentyear)) %>% 
  ungroup() %>%

@@

@@

# select data
d.selection <-
  d %>%
  filter(purpose == "advice") %>% 
  filter(!is.na(year))

# benchmark data
d.benchmarks <-
  d %>% 
  filter(benchmark == TRUE) %>% 
  distinct(stock, stockkeylabelold, assessmentyear, .keep_all = FALSE) %>% 
  mutate(benchmark = 1)

@@

  left_join(iAdvice, by=c("stockkeylabelold","assessmentyear", "purpose")) %>%
  select(-benchmark) %>% 
  mutate(assessmentmodel = tolower(assessmentmodel)) %>%

@@

  # determine stock definition change and model definition change
  arrange(stock, assessmentyear) %>% 
  group_by(stock) %>% 
  mutate(
    assessmentmodelchange = ifelse(assessmentmodel != lead(assessmentmodel), 2, 0),
    assessmentmodelchange = ifelse(assessmentmodel != lag(assessmentmodel) , 1, assessmentmodelchange),
    assessmentmodelchange = ifelse(assessmentmodel != lag(assessmentmodel) & 
                                   assessmentmodel != lead(assessmentmodel), 3, assessmentmodelchange),

    # add first and last year to stock definitions
    assessmentmodelchange = 
      case_when(
        assessmentyear == min(assessmentyear) & assessmentmodel != lead(assessmentmodel)  ~ 3, 
        assessmentyear == min(assessmentyear) & assessmentmodel == lead(assessmentmodel)  ~ 1,
        assessmentyear == max(assessmentyear) & assessmentmodel != lag(assessmentmodel) ~ 3, 
        assessmentyear == max(assessmentyear) & assessmentmodel == lag(assessmentmodel) ~ 2,
        TRUE ~ assessmentmodelchange
      ) 
    ) %>%

  # stock definition change
  mutate(
    stockdefinitionchange = ifelse(stockkeylabelold != lead(stockkeylabelold), 2, 0),
    stockdefinitionchange = ifelse(stockkeylabelold != lag(stockkeylabelold), 1, stockdefinitionchange),
    stockdefinitionchange = ifelse(is.na(stockdefinitionchange), 0, stockdefinitionchange),

    # add first and last year to stock definitions
    stockdefinitionchange = ifelse(assessmentyear == min(assessmentyear), 1, stockdefinitionchange),
    stockdefinitionchange = ifelse(assessmentyear == max(assessmentyear), 2, stockdefinitionchange)

  ) %>% 
  #     ifelse(assessmentyear == min(assessmentyear), 1, stockdefinitionchange),
  #   stockdefinitionchange = ifelse(assessmentyear == max(assessmentyear), 2, stockdefinitionchange)

  # benchmarks
  left_join(d.benchmarks, by=c("stock", "stockkeylabelold", "assessmentyear")) %>% 
  mutate(benchmark = ifelse(is.na(benchmark), 0, benchmark))

# dataset with stock definition changes
d.stockdefinitionchange <-
  d.assessments %>% 
  filter(stockdefinitionchange >= 1) %>% 
  distinct(stock, stockkeylabelold, assessmentyear, stockdefinitionchange, .keep_all = FALSE) %>%

  group_by(stock, stockkeylabelold) %>% 
  summarize(tmp = paste(assessmentyear, collapse="_")) %>% 
  separate(tmp, into=c("y1","y2"), sep="_") %>% 
  mutate_at(c("y1","y2"), funs(as.numeric)) %>% 
  arrange(stock, y1) %>%

  group_by(stock) %>% 
  mutate(rank = as.character(row_number()))

# dataset with assessment model periods
d.assessmentmodels3 <-
  d.assessments %>% 
  filter(assessmentmodelchange == 3) %>% 
  distinct(stock, assessmentyear, assessmentmodel, assessmentmodelchange, 
           .keep_all = FALSE) %>% 
  mutate(y1 = assessmentyear, y2 = assessmentyear+0.8) %>% 
  ungroup() %>% 
  select(-assessmentyear, -assessmentmodelchange)

d.assessmentmodels <-
  d.assessments %>% 
  filter(assessmentmodelchange %in% 1:2) %>% 
  distinct(stock, assessmentyear, assessmentmodel, assessmentmodelchange, 
           .keep_all = FALSE) %>%

  group_by(stock, assessmentmodel) %>% 
  summarize(tmp = paste(assessmentyear, collapse="_")) %>% 
  separate(tmp, into=c("y1","y2"), sep="_") %>% 
  mutate_at(c("y1","y2"), funs(as.numeric)) %>% 
  mutate(y2 = y2 +0.8) %>%

  # add the single year assessments
  bind_rows(d.assessmentmodels3) %>%

  arrange(stock, y1) %>%

  group_by(stock) %>% 
  mutate(rank = as.character(row_number()))

@@

  group_by(stock, variable, calctype) %>% 
  summarize(jonsson_ab   = mean(jonsson_ab),
            jonsson_asd  = sqrt(mean(jonsson_dev2)),
            mohn_rhoabs = mean(mohn_rhoabs),
            mohn_rho     = mean(mohn_rho), 
            firstyear    = min(year, na.rm=TRUE),
            lastyear     = max(year, na.rm=TRUE))

@@

  mutate(logB     = log(value),
         meanlogB = mean(log(value), na.rm=TRUE),
         logBdev  = logB - meanlogB,
         logBdev2 = (logBdev)^2,
         n        = n(),
         n_1      = n - 1) %>%

  # now do summing over years
  mutate(sumlogBdev2 = sum(logBdev2), na.rm=TRUE ) %>%

  # now summarize by year
  summarize(
    meanlogBdev = mean(logBdev, na.rm=TRUE),
    sumlogBdev2 = sum(logBdev2, na.rm=TRUE),
    n_1         = mean(n_1, na.rm=TRUE), 
    ralston_sigma = sqrt(sumlogBdev2/n_1)) %>%

@@

  group_by(stock, calctype) %>%
  distinct(stock, year, n_1, sumlogBdev2, .keep_all = FALSE) %>% 
  summarize(
    sumlogBdev2all = sum(sumlogBdev2),
    n_1            = sum(n_1), 
    firstyear    = min(year, na.rm=TRUE),
    lastyear     = max(year, na.rm=TRUE)
  ) %>%

@@

# data series with most recent (last) estimation
d.last <-
  d.pairs %>% 
  select(-assessmentyearprior) %>% 
  left_join(d.selection, by=c("stock", "variable", "assessmentyearlast" = "assessmentyear")) %>% 
  mutate(label="last") %>% 
  rename(assessmentyear = assessmentyearlast)

# data series with previous (prior) estimation
d.prior <-
  d.pairs %>% 
  select(-assessmentyearlast) %>% 
  left_join(d.selection, by=c("stock", "variable", "assessmentyearprior" = "assessmentyear")) %>% 
  mutate(label="prior") %>% 
  rename(assessmentyear = assessmentyearprior)

# pairwise data roughly
d.pairwise.roughly <-
  bind_rows(d.last, d.prior) %>% 
  group_by(stock, assessmentpair, variable) %>%

  # filter the appropriate period
  filter(year <= max(assessmentyear) - removelastyears,
         year >= max(assessmentyear) - removelastyears - numberofyears -5)

# pairwise data
d.pairwise <-
  d.pairwise.roughly %>% 
  group_by(stock, assessmentpair, variable) %>%

@@

  # filter only years that are in common in last and prior
  group_by(stock, assessmentpair, variable, year) %>% 
  mutate(n = n()) %>% 
  filter(n() == 2)

d.jumps <-
  d.pairwise %>%

@@

# jumps by year
d.jumps.byyear <-
  d.jumps %>% 
  group_by(variable, assessmentyear) %>% 
  summarise(n         = n(),
            n2        = sum(phi_abs > 0.1, na.rm=TRUE)) %>% 
  mutate(ratio = n2 / n )

@@

fig_nums(name    = "scalecomparisons", level = 1, display = FALSE, 
         caption = "Example of pairwise assessment comparisons for Northeast Atlantic mackerel  over a period of 10 years in common. The dotted lines indicate the average values of the different assessments which are used to calculate scale differences between assessments.")

fig_nums(name    = "scalejumps", level = 2, display = FALSE, 
         caption = "Retrospective scale difference (sigma) for 9 species derived from historical assessment records. Results from pairwise comparisons over periods of 10 years. Large dots  refer to years when the scale difference is larger than 10%. Colours indicate the decades. Black triangles indicate benchmark assessment years. The bottom part of the panels shows the assessment methods being deployed, the upper part of the panels the stock definitions.")

@@

fig_nums(name    = "jumpratio", level = 2, display = FALSE, 
         caption = "Ratio of scale-jumps in assessment relative to the number of assessment by year.")

getPalettePurples         <- colorRampPalette(brewer.pal(7, "Purples")[-(1:2)])
myPurples                 <- getPalettePurples(10)

getPaletteBlues         <- colorRampPalette(brewer.pal(7, "Blues")[-(1:2)])
myBlues                 <- getPaletteBlues(10)

@@

<!-- ################################################################################## -->
<!-- Plot pairwise comparisons                                                          -->
<!-- ################################################################################## -->

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

myvar     <- "stocksize"
myspecies <- "mac"
firstyear <- 1991

d.pairwise.means <-

  d.pairwise %>%

  filter(variable == myvar) %>% 
  filter(grepl(myspecies, stockkeylabelold)) %>% 
  filter(as.numeric(substr(assessmentpair,1,4)) >= firstyear) %>%

  # calculate means per period
  group_by(stock, assessmentpair, variable) %>% 
  mutate(   firstyear      = min(year),
            lastyear       = max(year)) %>% 
  group_by(stock, assessmentpair, variable, assessmentyear, label, firstyear, lastyear) %>% 
  summarise(value = mean(value, na.rm=TRUE)) %>% 
  gather(key=tmp, value=year, firstyear:lastyear) %>% 
  arrange(stock, assessmentpair, variable, assessmentyear, label)

d.pairwise %>%

  filter(variable == myvar) %>% 
  filter(grepl(myspecies, stockkeylabelold)) %>% 
  filter(as.numeric(substr(assessmentpair,1,4)) >= firstyear) %>%

  ggplot(aes(x=year, y=value, group=assessmentyear)) +
  theme_publication() +
  theme(
    panel.spacing.x = unit(1, "mm"),
    panel.spacing.y = unit(1, "mm"),
    strip.background = element_blank(),
    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(1,0,0,0, "mm")),
    strip.text.y     = element_text(angle = 0),
   # axis.text.x      = element_text(angle = 90, vjust = 0.5), 
    legend.position  = "none"
  ) +

  geom_line(data=filter(d.pairwise.roughly, 
                        variable == myvar, 
                        grepl(myspecies, stockkeylabelold),
                        as.numeric(substr(assessmentpair,1,4)) >= firstyear), size=0.5) +

  geom_line(aes(colour=factor(label)), size=1) +

  geom_line(data=d.pairwise.means, aes(x=year, y=value, colour=factor(label)), 
            linetype="dashed", lineend="butt") +
  geom_point(data=filter(d.pairwise.means, year==min(year)),
             aes(colour=factor(label)), size=1) +
  geom_point(data=filter(d.pairwise.means, year==max(year)),
             aes(colour=factor(label)), size=1) +

  scale_y_continuous(breaks = pretty_breaks(n=3), labels=scientific_format(digits=1)) +
  scale_x_continuous(breaks = pretty_breaks(n=2)) +
  expand_limits(y=0) +
  labs(x="", y="ssb (tonnes)") +
  facet_wrap( ~ assessmentpair, scales = "free_x", ncol=5)

*`r fig_nums("scalecomparisons")`*

@@

  ggplot(aes(x=assessmentyear, y=value, group=stock)) +
  theme_publication() +
  theme(
    panel.spacing    = unit(0.1, "lines"),
    panel.grid.major.x = element_line(colour = "gray80"),
    strip.background = element_blank(),
    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), angle = 180),
    strip.placement = "outside",
    legend.position  = "none"
  ) +

  geom_rect(data=filter(d.stockdefinitionchange),
            aes(xmin = y1, xmax=y2, ymin = -1.0, ymax=1.0, fill=rank), 
            alpha=0.4, colour=myPurples[6], inherit.aes = FALSE) +

  geom_dl(data=filter(d.stockdefinitionchange),
          aes(label  = stockkeylabelold, x = y1, y = -0.95),
          method = list(dl.combine("first.points"), rot = 90, cex = 0.6, hjust=0.0, vjust=0.9),
          inherit.aes = FALSE, colour=brewer.pal(6, "Purples")[6]) +

@@

@@

  geom_point(data=filter(d.assessments, benchmark == 1),
            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +

  geom_point(data=filter(d.assessments, assessmentmodelchange == 1),
            aes(x=assessmentyear-0.1, y=-1.0), colour="red", fill="red", shape=24, size=2) +

  coord_cartesian(ylim=c(-1, 1),
                  xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded

  scale_fill_manual(values = brewer.pal(6, "Purples")[-(1:2)]) +

  scale_y_continuous(breaks = pretty_breaks(), labels=percent_format()) +
  scale_x_continuous(breaks = pretty_breaks()) +

  labs(x=" ", y="") +
  facet_wrap(~stock)

@@

# plot scale jumps between assessments vs assessment methods

d.jumps %>% 
  distinct(stock, variable, assessmentyear, phi) %>%
  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
  gather(key=metric, value=value, c(phi)) %>%

  ggplot(aes(x=assessmentyear, y=value, group=stock)) +
  theme_publication() +
  theme(
    panel.spacing    = unit(0.1, "lines"),
    panel.grid.major.x = element_line(colour = "gray80"),
    strip.background = element_blank(),
    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), angle = 180),
    strip.placement = "outside",
    legend.position  = "none"
  ) +

  geom_rect(data=filter(d.assessmentmodels),
            aes(xmin = y1, xmax=y2, ymin = -1.0, ymax=1.0, fill=rank), 
            alpha=0.4, colour=myBlues[6], inherit.aes = FALSE) +

  geom_dl(data=filter(d.assessmentmodels),
          aes(label  = assessmentmodel, x = y1, y = -0.95),
          method = list(dl.combine("first.points"), rot = 90, cex = 0.6, hjust=0.0, vjust=0.9),
          inherit.aes = FALSE, colour=myBlues[6]) +

  geom_point(data=filter(d.assessments, benchmark == 1),
            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +

  coord_cartesian(ylim=c(-1, 1),
                  xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded

  scale_fill_manual(values = myBlues) +

  scale_y_continuous(breaks = pretty_breaks(), labels=percent_format()) +
  scale_x_continuous(breaks = pretty_breaks()) +

  labs(x=" ", y="") +
  facet_wrap(~stock)

# plot scale jumps between assessments vs assessment methods

d.jumps %>% 
  distinct(stock, variable, assessmentyear, phi) %>%
  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
  gather(key=metric, value=value, c(phi)) %>%

  ggplot(aes(x=assessmentyear, y=value, group=stock)) +
  theme_publication() +
  theme(
    panel.spacing    = unit(0.1, "lines"),
    panel.grid.major.x = element_line(colour = "gray80"),
    strip.background = element_blank(),
    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), angle = 180),
    strip.placement = "outside",
    legend.position  = "none"
  ) +

  geom_rect(data=filter(d.stockdefinitionchange),
            aes(xmin = y1, xmax=y2, ymin = 0.00, ymax=1.0), 
            alpha=0.4, fill=myPurples[6], colour=myPurples[6], inherit.aes = FALSE) +

  geom_dl(data=filter(d.stockdefinitionchange),
          aes(label  = stockkeylabelold, x = y1, y = 0.95),
          method = list(dl.combine("first.points"), rot = 90, cex = 0.6, hjust=1.0, vjust=0.9),
          inherit.aes = FALSE, colour=brewer.pal(6, "Purples")[6]) +

  geom_rect(data=filter(d.assessmentmodels),
            aes(xmin = y1, xmax=y2, ymin = -1.0, ymax=0.0, fill=rank), 
            alpha=0.4, colour=myBlues[6], inherit.aes = FALSE) +

  geom_dl(data=filter(d.assessmentmodels),
          aes(label  = assessmentmodel, x = y1, y = -0.95),
          method = list(dl.combine("first.points"), rot = 90, cex = 0.6, hjust=0.0, vjust=0.9),
          inherit.aes = FALSE, colour=myBlues[6]) +

  geom_point(data=filter(d.assessments, benchmark == 1),
            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +

  coord_cartesian(ylim=c(-1, 1),
                  xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded

  scale_fill_manual(values = myBlues) +

  scale_y_continuous(breaks = pretty_breaks(), labels=percent_format()) +
  scale_x_continuous(breaks = pretty_breaks()) +

  labs(x=" ", y="") +
  facet_wrap(~stock)

@@

<!-- ################################################################################## -->
<!-- COmpare the number of scale jumps (as ratio) over time                             -->
<!-- ################################################################################## -->

```{r jumpratio, echo=FALSE, fig.asp=0.5, fig.align="center", message=FALSE, warning=FALSE}

d.jumps.byyear %>% 
  filter(assessmentyear >= 1990) %>%

  ggplot(aes(x=assessmentyear, y=ratio)) +
  theme_publication() +
  theme(
    panel.spacing    = unit(0.1, "lines"),
    panel.grid.major.x = element_line(colour = "gray80"),
    strip.background = element_blank(),
    strip.text.x     = element_text(size=8, face="bold", hjust=1, margin = margin(2,0,2,0, "mm")),
    strip.placement = "outside",
    legend.title    = element_blank(),
    legend.position  = "none"
  ) +
  geom_bar(stat="identity") +
  labs(x="", y="ratio")

*`r fig_nums("jumpratio")`*
